---
title: "Simulating_data"
output: html_document
date: "2025-04-01"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(broom)
library(tidyverse)
```

# Propensity Score Matching

## Simulating the data

### Let's simulate some Amazon user demographic data

To do this we first need some data. Let's simulate some data to illustrate propensity score matching. Let's pretend that we work for Amazon. We are going to "treat" some individuals with free-shipping to see what happens to their average purchase.

```{r}
sample_size <- 10000
set.seed(853)

# Variables

unique_person_id = c(1:sample_size)
age = runif(n = sample_size, min = 18, max = 100)
city = sample(x = c("Toronto", "Montreal", "Calgary"),
      size = sample_size,
      replace = TRUE)
gender = sample(x = c("Female", "Male", "Other/decline"),
      size = sample_size,
      replace = TRUE,
      prob = c(0.49, 0.47, 0.02))
income = rlnorm(n = sample_size,
                    meanlog = 0.5, 
                    sdlog = 1)

# Data Frame

amazon_purchase_data <- tibble(unique_person_id,
                               age,
                               city,
                               gender,
                               income)

```


### Adding in the treatment (free shipping)

Now we need to add some probability of being treated with free shipping, which depends on our variables. Let's assume younger, higher-income, male and in Toronto all make it slightly more likely.

Let's start by creating a scoring system.

```{r}
amazon_purchase_data <-
  amazon_purchase_data %>% 
  mutate(age_num = case_when(
    age < 30 ~ 3,
    age < 50 ~ 2,
    age < 70 ~ 1,
    TRUE ~ 0),
    city_num = case_when(
      city == "Toronto" ~ 3,
      city == "Montreal" ~ 2,
      city == "Calgary" ~ 1,
      TRUE ~ 0),
    gender_num = case_when(
      gender == "Male" ~ 3,
      gender == "Female" ~ 2,
      gender == "Other/decline" ~ 1,
      TRUE ~ 0),
    income_num = case_when(
      income > 3 ~ 3,
      income > 2 ~ 2,
      income > 1 ~ 1,
      TRUE ~ 0)
  )

```

Now that we have the scoring system to help us "indicate" whether or not someone is likely to get free shipping, let's actually create the free shipping (treatment) indicator variable.

```{r}
amazon_purchase_data <- 
  amazon_purchase_data %>% 
  rowwise() %>% 
  mutate(sum_num = sum(age_num, city_num, gender_num, income_num)) %>% 
  mutate(softmax_prob = exp(sum_num)/exp(12)) %>% 
  mutate(free_shipping = sample(
           x = c(0:1),
           size = 1,
           replace = TRUE,
           prob = c(1-softmax_prob, softmax_prob)
         ) 
      ) %>% 
  ungroup()
```

Great, now we have a treatment variable, but the `age_num`, `gender_num`, etc., are somewhat distracting.


### Cleaning the simulated data a bit

Okay, now that we finally have the treatment variable, we should remove all the variables we simulated to get us there.

```{r}
amazon_purchase_data <-
  amazon_purchase_data %>% 
  dplyr::select(-age_num, -city_num, -gender_num, -income_num, 
                -sum_num, -softmax_prob)

```


### Adding in the outcome (spending)

Finally, we need to have some measure of a person's  average spend. We want those with free shipping to be slightly higher than those without.

```{r}
amazon_purchase_data <-
  amazon_purchase_data %>% 
  mutate(mean_spend = if_else(free_shipping == 1, 60, 50)) %>% 
  rowwise() %>% 
  mutate(average_spend = rnorm(1, mean_spend, sd = 5)
  ) %>% 
  ungroup()
```

Great, now I will just remove the mean spending variable.

```{r}
amazon_purchase_data <- amazon_purchase_data %>% select(-mean_spend)
```

### Cleaning a bit more

Fix the class on some variables.

```{r}
amazon_purchase_data <-
  amazon_purchase_data %>% 
  mutate(city = as.factor(city), 
         gender = as.factor(gender), 
         free_shipping = as.factor(free_shipping)) # Change some to factors

table(amazon_purchase_data$free_shipping)
```

## Saving the data

```{r}
write_csv(amazon_purchase_data, "Amazon.csv")
```

